Prefetching SPARQL query cacher

### 2016-08-03 dissertation/cache-proxy.tex
Martins final comments

Following this discussion, consider this example:
\begin{example}{A four-triple pattern BGP}\label{ex:bgp4}
\begin{verbatim}
?s <http://example.org/m/p> "1" ;
   <http://example.org/m/q> <http://example.org/m/a> ;
   <http://example.org/m/p> ?o .
?o <http://example.org/m/b> "2" .
\end{verbatim}
\end{example}
Say that the first triple pattern has a result in the cache. This will
not cause Problem~\ref{prob:cachecartesian} to occur (it would have
occurred if the third triple pattern had a cached result, however), so
we can safely use the cache. The results of the three remaining triple
patterns must be fetched from a remote server. There are essentially
two reasonable choices: Either fetch the result from an LDF server as
argued by \cite{verborgh2014querying} and join them locally or send a
single, rewritten SPARQL query to the remote endpoint for the
evaluation. When the results are returned from both the cache and the
remote endpoint, they must then finally be joined to produce the final
result returned to the client. A string representation of the plan
tree from the current implementation reflects this and reads (note
that accessing the cache is represented by an iterator plan):
\begin{example}{An example Attean plan with a remotely executed BGP}
\begin{verbatim}
- Hash Join { s } (distinct cost: 74)
-   SPARQLBGP (cost: 60)
-     Quad { ?s, <http://example.org/m/p>, ?o, <http://test.invalid/graph> }
-     Quad { ?o, <http://example.org/m/b>, "2", <http://test.invalid/graph> }
-     Quad { ?s, <http://example.org/m/q>, <http://example.org/m/a>, <http://test.invalid/graph> }
-   Iterator (?s with 2 elements) (cost: 2)
\end{verbatim}
\end{example}

Merging the quad plans into a single SPARQL BGP plan is not
straightforward in the presence of a quad pattern that is to be looked
up in e.g. a cache. If we assign the letters $A$, $B$, $C$ and $D$ to
quad patterns \todo{correctly expressed?} in the order in
Example~\ref{ex:bgp4}, the initial join tree produced by parsing may
be $ ((A \bowtie B )\bowtie C) \bowtie D) $. Now, we are interested in
a plan on the form $A \bowtie BCD$, i.e. $BCD$ is reduced to a more
fundamental plan type, in this example a SPARQL BGP. This will
minimise the number of HTTP requests, in much the same way as the
exclusive groups of
\cite{springerlink:10.1007/978-3-642-25073-6-38}. To facilitate this,
we employ \emph{join rotation}. A simple example of join rotation
would be this: We first generate subplans for both $((A \bowtie B
)\bowtie C) \bowtie D) $ and $ (A \bowtie B )\bowtie (C \bowtie D)
$. Then, $C$ and $D$ can be merged to a SPARQL BGP to $ (A \bowtie B
)\bowtie CD $. Another rotation can then be performed: $ A \bowtie (B
\bowtie CD) $ and then coalesced to the final result. The planner will
generate alternative plans for both rotated and unrotated plans, and
all generated plans are subject to cost estimation, so whether a large
BGP plan will be evaluated remotely, or as a join of local cached
results and a smaller remote BGP, or a remote Triple Pattern Fragments
server, is up to the cost estimation. This will be discussed in detail
in Section~\ref{sec:costheuristics}.

>>>>

Following this discussion, consider this example:
\begin{example}{A four-triple pattern BGP}\label{ex:bgp4}
\begin{verbatim}
?s <http://example.org/m/p> "1" .                      # A
?s <http://example.org/m/q> <http://example.org/m/a> . # B
?s <http://example.org/m/p> ?o .                       # C
?o <http://example.org/m/b> "2" .                      # D
\end{verbatim}
\end{example}
Say that the first triple pattern has a result in the cache. This will
not cause Problem~\ref{prob:cachecartesian} to occur (it would have
occurred if the third triple pattern had a cached result, however), so
we can safely use the cache. The results of the three remaining triple
patterns must be fetched from a remote server. There are essentially
two reasonable choices: Either fetch the result from an LDF server as
argued by \cite{verborgh2014querying} and join them locally or send a
single, rewritten SPARQL query to the remote endpoint for the
evaluation. When the results are returned from both the cache and the
remote endpoint, they must then finally be joined to produce the final
result returned to the client. A string representation of the plan
tree from the current implementation reflects this and reads (note
that accessing the cache is represented by an iterator plan):
\begin{example}{An example Attean plan with a remotely executed BGP}
\begin{verbatim}
- Hash Join { s } (distinct cost: 74)
-   SPARQLBGP (cost: 60)
-     Quad { ?s, <http://example.org/m/p>, ?o, <http://test.invalid/graph> }
-     Quad { ?o, <http://example.org/m/b>, "2", <http://test.invalid/graph> }
-     Quad { ?s, <http://example.org/m/q>, <http://example.org/m/a>, <http://test.invalid/graph> }
-   Iterator (?s with 2 elements) (cost: 2)
\end{verbatim}
\end{example}

Merging the quad plans into a single SPARQL BGP plan is not
straightforward in the presence of a quad pattern that is to be looked
up in e.g. a cache. If we assign the letters $A$, $B$, $C$ and $D$ to
triple patterns in the order expressed in
Example~\ref{ex:bgp4}, the initial join tree produced by parsing may
be $ (((A \bowtie B )\bowtie C) \bowtie D) $. Now, we are interested in
a plan on the form $A \bowtie BCD$, i.e. $BCD$ is reduced to a more
fundamental plan type, in this example a SPARQL BGP. This will
minimise the number of HTTP requests, in much the same way as the
exclusive groups of
\cite{springerlink:10.1007/978-3-642-25073-6-38}. To facilitate this,
we employ \emph{join rotation}. A simple example of join rotation
would be this: We first generate subplans for both $(((A \bowtie B
)\bowtie C) \bowtie D) $ and $ (A \bowtie B )\bowtie (C \bowtie D)
$. Then, $C$ and $D$ can be merged to a SPARQL BGP to $ (A \bowtie B
)\bowtie CD $. Another rotation can then be performed: $ A \bowtie (B
\bowtie CD) $ and then coalesced to the final result. The planner will
generate alternative plans for both rotated and unrotated plans, and
all generated plans are subject to cost estimation, so whether a large
BGP plan will be evaluated remotely, or as a join of local cached
results and a smaller remote BGP, or a remote Triple Pattern Fragments
server, is up to the cost estimation. This will be discussed in detail
in Section~\ref{sec:costheuristics}.
